\normalfont\spacedallcaps{}

### 2019-03-26 Document/Proposa.tex
nnn

\input{structure.tex}
\usepackage{cite}
\usepackage{adjustbox}
\usepackage{booktabs}
\usepackage{float}
\hyphenation{Fortran hy-phen-ation} 
\captionsetup{font=footnotesize}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{tabularx, makecell}%
\renewcommand\theadfont{\normalsize\bfseries}
        \usepackage{etoolbox} %
        \AtBeginEnvironment{tabularx}{\setlist[enumerate, 1]{wide, leftmargin=*, itemsep=0pt, before=\vspace{-\dimexpr\baselineskip +2 \partopsep}, after=\vspace{-\baselineskip}}}

@@

\subsection{\textbf{tRNA identity detereminant visualization}}
There are 21 (excluding initiators) class of tRNA genes, which code for 21 class of amino acids. We reperesent set of 22 single letter amino acid code as $Y = \{A, C, D, E, F, G, H, I, J, K, L, M, N, P, Q, R, S, T, V, W, X, Y \}$. tRNAs with different anticodon that accept same amino acid are called isoacceptors. Here we use IUPAC one-letter code of the amino acid to label tRNAs. letter X is used for labeling initiators. tRNA identity refers to the specific amino acid it codes for. Enzymes called aminoacyl-tRNA synthetases (aaRSs) bind the right amino acid with the right tRNA. each aaRS recognizes the right tRNA based on its sequence and structural features which are called tRNA identity determinants. Bellow I describe four identity determinants visualization tools we use in out research. 
\subsubsection*{\textbf{Function Logo}}
Function logos \cite{Freyhult2006VisualizingBT} are a generalization of sequence logos \cite{Schneider} made from a set of aligned tRNA sequences with length L and are defined as the Cartesian product of a state $x \in X$ where $X = \{A, C, G, U, - \}$ and $l \in L$ where $1 \leqslant l \leqslant L$. Every specific state $x$ at position $l$ is called feature $x_l$. Functional information $I_l(Y|x)$ that a state x confers about the frequencies of different classes $Y$ at position $l$ is calculated for each feature $x_l$ as  $I_l(Y|x)=H(Y)-e(n_l(x))-H_l(Y|x))$. Where $H_l(Y|x)=-\sum_{y \in Y} p_l(y|x)\log_2(p_l(y|x))$ is the class entropy calculated based on the frequency of sequences that carry state $x$ at position $l$, $H(Y)$ is the background class entropy which is calculated as $-\sum_{y \in Y} p(y)\log_2(p(y))$, and $e(n_l(x))$ is a correction factor to correct for biases caused by small sample size. \\
Function logo is a symbol-stacked-bar graph with positions on $x$ axis and informations on $y$ axis where each symbol within a bar is one of the functional classes of tRNAs in set $Y$ and symbols are sorted based on their height. Height of each symbol $y$ ($y \in Y$) for feature $x_l$ is proportional to the frequency of sequences of class $y$ with that feature and is calculated as $(\sum_{w \in Y} p_l(w|x)/p(w)) I_l(Y|x)$. Figure~\ref{fig:Flogo} shows an example of function logo for state C generated by tsfm. Information at each bar are shown in bits and the maximum information of each stack is 4.2 bits. Postition 22 is an example of an indentity determinant which is fixed for functional class S. This means that sequences with C at position 22 are very likely to belong to functional class Serine.

@@

\subsubsection*{Detecting possible anticodon shifts}
Anticodon is one of the major identity determinant in tRNAs. It is possible for a tRNA gene to switch to a different functional class after one or more mutations in its anticodon sequence and it has been shown to be possible in vitro as well (Schulman and Pelka 1989; Pallanck and Schulman 1991). mutations of bases in anticodon sequence resulting in change of tRNA’s amino acid charging is called alloaccepter shifts. relative number of anticodon shifts in a taxa can be the can suggest that tRNA gene redundancy is likely the driving factor (ref?)
also, by detecting the antocodon shifts, we can find sites with high covariation with anticodon shifts as potential determinants of tRNAs.\\ 
Previous works on anticodon shifts in eukaryotic tRNA genes (ref) has been done using a synteny-conservation-based method which looks for different anticodons within ortholog tRNAs as potential anticodon shifts. although this method is .... Mappings of flanking regions for each tRNA agains all other flanking regions and and ortholog set compilation may not be computationally efficient. By building a tRNA classifer we can score all the query tRNAs against all the models with time complexity of length of query. further, tRNA gene models with mismatch identity assigned by the classifier and the identity assigned by genefinders based on anti-codon, can be used as the potential anticodon shifts. these quences can be further be used in flank-mapping method for the perpuse of verification.

\subsubsection*{Detecting differences of indentity determinants between Trypanosoma and Human tRNA genes} 
The trypanosomatids are flagellated protozoan parasite that include species Trypanosoma brucei, Trypanosoma cruzi and Leishmania major also known as TriTryp genomes. Three major diseases caused by these species are African trypanosomiasis,  South American trypanosomiasis and leishmaniasis in order, which can cause permanent disability or death in humans. Trypanosomatids are eukaryotic single cells and it is a challenge to develop drugs which can selectively target their pathogen and not affecting human host. Available treatments for Leishmanianis used for many years has been found to evolve drug resistance (3 ref). It is important to find targets that has been diverged significantly from human with less chance of developing resistance. Previous works has shown that targeting  aaRSs via its interaction with tRNA can be a great target for parasites(22). Here, we would like to reannotate and classify tRNA genes in Trypanosomatids, study evolution of these tRNAs across TriTyp genomes and find differences of tRNA identity determinants between TriTryps and humans to be used in future as a target to develop non-toxic drugs for human.

% write the significance of each aim
%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHYyour own dat
%----------------------------------------------------------------------------------------
\newpage
\renewcommand{\refname}{\spacedlowsmallcaps{References}} % For modifying the bibliography heading

>>>>

\input{structure.tex}
\usepackage[a4paper, total={5.8in, 10in}]{geometry}
\usepackage{cite}
\usepackage{adjustbox}
\usepackage{booktabs}
\usepackage{float}
\hyphenation{Fortran hy-phen-ation} 
\captionsetup{font=footnotesize}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{tabularx, makecell}%
\renewcommand\theadfont{\normalsize\bfseries}
        \usepackage{etoolbox} %
        \AtBeginEnvironment{tabularx}{\setlist[enumerate, 1]{wide, leftmargin=*, itemsep=0pt, before=\vspace{-\dimexpr\baselineskip +2 \partopsep}, after=\vspace{-\baselineskip}}}

@@

\subsection{\textbf{tRNA identity detereminant visualization}}
There are 21 (excluding initiators) class of tRNA genes, which code for 21 class of amino acids. We reperesent set of 22 single letter amino acid code as $Y =$ \{A, C, D, E, F, G, H, I, J, K, L, M, N, P, Q, R, S, T, V, W, X, Y \}. tRNAs with different anticodon that accept same amino acid are called isoacceptors. Here we use IUPAC one-letter code of the amino acid to label tRNAs. letter X is used for labeling initiators. tRNA identity refers to the specific amino acid it codes for. Enzymes called aminoacyl-tRNA synthetases (aaRSs) bind the right amino acid with the right tRNA. each aaRS recognizes the right tRNA based on its sequence and structural features which are called tRNA identity determinants. Bellow I describe four identity determinants visualization tools we use in out research. 
\subsubsection*{\textbf{Function Logo}}
Function logos \cite{Freyhult2006VisualizingBT} are a generalization of sequence logos \cite{Schneider} made from a set of aligned tRNA sequences with length L and are defined as the Cartesian product of a state $x \in X$ where $X = \{A, C, G, U, - \}$ and $l \in L$ where $1 \leqslant l \leqslant L$. Every specific state $x$ at position $l$ is called feature $x_l$. Functional information $I_l(Y|x)$ that a state x confers about the frequencies of different classes $Y$ at position $l$ is calculated for each feature $x_l$ as  $I_l(Y|x)=H(Y)-e(n_l(x))-H_l(Y|x))$. Where $H_l(Y|x)=-\sum_{y \in Y} p_l(y|x)\log_2(p_l(y|x))$ is the class entropy calculated based on the frequency of sequences that carry state $x$ at position $l$, $H(Y)$ is the background class entropy which is calculated as $-\sum_{y \in Y} p(y)\log_2(p(y))$, and $e(n_l(x))$ is a correction factor to correct for biases caused by small sample size. \\
Function logo is a symbol-stacked-bar graph with positions on $x$ axis and informations on $y$ axis where each symbol within a bar is one of the functional classes of tRNAs in set $Y$ and symbols are sorted based on their height. Height of each symbol $y$ ($y \in Y$) for feature $x_l$ is proportional to the frequency of sequences of class $y$ with that feature and is calculated as $(\sum_{w \in Y} p_l(w|x)/p(w)) I_l(Y|x)$. Figure~\ref{fig:Flogo} shows an example of function logo for state C generated by tsfm. Information at each bar are shown in bits and the maximum information of each stack is 4.2 bits. Postition 22 is an example of an indentity determinant which is fixed for functional class S. This means that sequences with C at position 22 are very likely to belong to functional class Serine.

@@

\subsubsection*{tRNA anticodon shifts}
Anticodon is one of the major identity determinant in tRNAs. It is shown in vitro that it is possible for a tRNA gene to switch to a different functional class after one or more mutations in its anticodon sequence \cite{shifts1,shifts2}. Mutation in anticodon resulting in change of tRNA’s amino acid charging is called alloaccepter shift. Previous work on detecting anticodon mutations in eukaryotic genomes suggests that tRNA redundancy can be a reason for relative number of anticodon shifts in a taxa \cite{Rogers2014tRNAAS}. A synteny-conservation-based method is used in this work which looks for different anticodons within ortholog tRNAs as potential anticodon shifts. Although the restriction of a tRNA gene having at least one ortholog to be considered active is important, mapping of flanking regions for each tRNA against all other flanking regions to find ortholog sets may not be computationally efficient. A tRNA classifier can score all the query tRNAs against all the profile models of a classifier with time complexity of length of queries. tRNA gene models with mismatch identity assigned by the classifier and the identity assigned by gene finders based on anticodon, can be used as the potential anticodon shifts and be used in flank-mapping method for the purpose of verification. Further, by predicting anticodon shifts, we can find sites that covary with these anticodon mutations as potential determinants of tRNAs.

\subsubsection*{Detecting differences of indentity determinants between Trypanosoma and Human tRNA genes} 
The trypanosomatids are flagellated protozoan parasite that include species Trypanosoma brucei, Trypanosoma cruzi and Leishmania major also known as TriTryp genomes. Three major diseases caused by these species are African trypanosomiasis,  South American trypanosomiasis and leishmaniasis in order, which can cause permanent disability or death in humans. Trypanosomatids are eukaryotic single cells and it is a challenge to develop drugs which can selectively target their pathogen without affecting human host. Available treatments for Leishmanianis used for many years has been found to evolve drug resistance \cite{Resistance1,Resistance2}. It is important to find targets that has been diverged significantly from human with less chance of developing resistance. Previous works has shown that targeting  aaRSs via its interaction with tRNA can be a great target for parasites \cite{aaRS}. Here, we would like to re annotate and classify tRNA genes in Trypanosomatids, study their evolution across TriTryp genomes and find differences of tRNA identity determinants between TriTryps and humans to be used in future as a target to develop non-toxic drugs for human. 
% write the significance of each aim
%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHYyour own dat
%----------------------------------------------------------------------------------------
\newpage
\renewcommand{\refname}{\spacedlowsmallcaps{References}} % For modifying the bibliography heading
